%\section{Galois Extensions and Galois Correspondence}

 \subsection{Galois Extensions and Galois Correspondence}


We are now ready to present a special kind of field extension:

\begin{definition}
    A finite field extension is \textit{Galois} if it is both normal and separable. 
\end{definition}

By Theorem \ref{thm:separable-extension-in-C}, for a finite field extension in $\C$, being `Galois' is a synonym of being `normal', as its separability is automatic. 
We now give some further properties for Galois extensions in general fields. The statements in the following theorem (which formalises ideas in \cite[Lecture~8]{galois-theory-lectures}) should not seem surprising, as we have mentioned (ii) and (iii) in the previous section and (iv) is apparently based on the definitions of normal and separable extensions. 


\begin{theorem} \label{thm:fixed}
	Let $L/K$ be a finite field extension, and let $G = \Gal(L/K)$. Then the following statements are equivalent:
 	(i) $L/K$ is a Galois extension;
	  (ii) $[L:K] = |G|$;
	  (iii) $K = \Fix(G)$;
	  (iv) $L$ is a splitting field of a separable polynomial over $K$.
	% \begin{enumerate}[label=(\roman*)]
	% 	\item $L/K$ is a Galois extension;
	% 	\item $[L:K] = |G|$;
	% 	\item $K = \Fix(G)$;
	% 	\item $L$ is a splitting field of a separable polynomial over $K$;
	% 	%        \item There is a natural bijection between the intermediate fields $M$ such that $K \subseteq M \subseteq L$ and the subgroups $H \le G$, where $\alpha$ and $\beta$ defined by $\alpha(M) =  \Gal (L/M)$ and $\beta(H) =  \Fix(H)$ are inverses of each other.
	% \end{enumerate}
	% Given a finite Galois extension $L/K$ and its Galois group $G=\Gal(L/K)$, we have $K = \Fix(G)$.
\end{theorem}

%The proof has two parts: the first part establishes the equivalence of the first four statements and the second part works on the equivalence of the fifth and the rest. 

%\begin{proof}[Proof, first part]
\begin{proof}
	
	%	We now prove the equivalence of the first four statements. We will work on the fifth statement separately shortly afterwards.
	
	
	(i) $\Rightarrow$ (ii). Consider a $K$-monomorphism $\phi: K \to \overline{K}$. We now extend $\phi$ to a $K$-monomorphism $\phi':L \to \overline{K}$ in the same pattern as in the proof of Lemma \ref{thm:galois-group-order-upper-bound}. However, at each step, we claim that the number of ways to extend is equal to $[K_i : K_{i-1}]$. Indeed, $\alpha_i$ is separable and hence $f_i$ has exactly $[K_i : K_{i-1}]$ distinct roots, and $\overline K$ contains all these roots. Therefore, combining all the steps gives us in total $[L:K]$ ways to extend $\phi$ to $\phi'$. Also, since $L/K$ is normal, by Theorem \ref{thm:normal-equiv-def}, each $K$-monomorphism $\phi': L \to \overline{K}$ is a $K$-automorphism of $L$, and thus each $\phi'$ corresponds to an element in $\Gal(L/K)$. Hence $[L:K] = \Gal(L/K)$. 
	
	(ii) $\Rightarrow$ (iii). Define $F:= \Fix(G)$. We have $K \subseteq F \subseteq G$. Then by assumption, we have $[L:K] = |G|$; each $K$-automorphism of $L$ in $G$ is also a $F$-automorphism of $L$ in $\Gal(L/F)$, and thus $G \subseteq \Gal(L / F)$ and $|G| \le |\Gal(L / F)|$; by Theorem \ref{thm:galois-group-order-upper-bound}, $|\Gal(L/F)| \le [L:F]$; by Theorem \ref{thm:tower-theorem}, we see that $ [L:F] \le [L:F][F:K] = [L:K]$. Combining them gives $
	[L:K] = |G| \le |\Gal(L/F)| \le  [L:F] \le [L:K],
	$
	and thus all the $\le$ signs must take equality.  
	Thus $[L:K]=[L:F]$ and therefore by Theorem \ref{thm:tower-theorem}, we see that $[F:K]$ must equal $1$, and thus $F:=\Fix(G) = K$.
	
	(iii) $\Rightarrow$ (iv). Let $\alpha \in L$. Consider $\Orb_G(\alpha) \subseteq L$, the orbit of $\alpha$ under the $G$-action on $L$. Then since $G$ is finite, $\Orb_G(\alpha) = \{ \alpha_1 = \alpha, \alpha_2, \ldots, \alpha_n\}$. Consider the polynomial $f(t) = (t-\alpha_1) \ldots (t-\alpha_n)$ over $L$.  $f$ is separable in $L$, as $\alpha_i$ are distinct. Also, any $\sigma \in G$ acting on $\alpha_1, \dots, \alpha_n$ only permutes the elements, and thus $\sigma$ fixes $f$. Then the coefficients of $f$ must be in $\Fix(G)$, but $\Fix(G) = K$ by assumption. Hence $f$ is over $K$. Clearly, $L$ is a splitting field of $f$ over $K$. 
	
	(iv) $\Rightarrow$ (i) is trivial.
\end{proof}


\begin{corollary} \label{thm:galois-intermediate}
	If $L/K$ is a Galois extension, then for an intermediate field $M$ such that $K \subseteq M \subseteq L$, $L/M$ is also a Galois extension. 
\end{corollary}
\begin{proof}
	By (iv) of Theorem \ref{thm:fixed}, $L$ is the splitting field of some polynomial $f$ over $K$, but $f$ is also a polynomial over $M$. Hence $L/M$ is Galois.
\end{proof}

\begin{example}
	Note that in the above corollary, $L/K$ being a Galois extension does not imply that $M/K$ is Galois. This is in general false; consider for example $K = \Q, M = \Q(\sqrt[3]{2})$ and $L = \Q(\sqrt[3]{2}, \zeta)$, where $\zeta = e^{2\pi i / 3}$. Then $L/K$ and $L/M$ are Galois but $M/K$ is not. 
\end{example}

The most important property of Galois extensions is that of Galois correspondence, which we now establish. The following is needed for the main theorem. 


%\begin{theorem}
%	Let $L$ be a field and let $G \le \Aut(L)$ be finite. Let $K = \Fix(G)$. Then $[L:K] =|G|$. 
%\end{theorem}
%
%\begin{proof}
%	\TODO(10.5)
%\end{proof}




% \subsection{Fundamental Theorem of Galois Theory}

%\begin{theorem}
%	Let $M$ be a field and let $G \le \Aut(M)$. Then $M / \Fix(G)$ is a Galois extension.
%\end{theorem}
%
%\begin{proof}
%	\TODO
%\end{proof}
%
%\begin{theorem}
%    Let $G$ be a finite subgroup of the group of automorphisms of a field $L$, and let $K = \Fix(G)$. Then $[L : K] = |G|$. 
%\end{theorem}
%
%\begin{proof}
%    \TODO
%\end{proof}

% \noindent We now look at a theorem which helps to characterise and identify fields, extensions and Galois groups.


\begin{theorem}\label{thm:equal-subgroup-fixed-points}
    Let $L/K$ be a finite Galois extension, with $G=\Gal(L/K)$, then if we have a subgroup $H\leq G$ with $\Fix(H)=K$ then $H=G$.
\end{theorem}

\begin{proof}
    Suppose $H\leq G$ with $\Fix(H)=K$. Then, let $L = K(a)$, then $\Orb_H(a)$ denotes the orbit under the group action of $H$ in $L$, and we can assume that $\Orb_H(a) = \{a_1,...,a_m\}.$ From this we can see that $|\Orb_H(a)| = m$, and let $[L:K]=n$.
    Then we know that the $m$ cannot be larger than the size of $H$, and since $H\leq G \implies |H|\leq|G|$, we get $|\Orb_H(a)| = m \leq |H| \leq |G| = [L:K] = n$.
    
   
    On the other hand, by evaluating the elementary symmetric polynomials at each of the elements in $\Orb_H(a)$, we get that, by the fact they are symmetric, these are just the elements in $\Fix(H)=K$. Therefore, we see that the coefficients in the polynomial: $f(t) := (t-a_1)(t-a_2)...(t-a_m)$ lies in $K$, and we have $f(a)=0$. Then, our minimal polynomial $\mu_a(t)$ divides $f(t)$ and $\partial f \geq \partial \mu_a(t)$ and hence $m \geq n$. Hence we have $m \geq n$ and $n \geq m$, so therefore we have that $m=n$, which tells us $|H|=|G|$ and so $H=G.$
\end{proof}

We are now ready to present the core result, the Galois correspondence of Galois extensions, which is a bijection between its intermediate fields and the subgroups of its Galois groups. 

\begin{theorem}[The Fundamental Theorem of Galois Theory] \label{thm:fundamental-theorem} Given a Galois extension $L/K$ and its Galois group $G = \Gal(L/K)$, there is a natural bijection between subgroups $H\leq G$ and the intermediate fields $M$ such that $K \subseteq M \subseteq L$. Define
 $\alpha:M \mapsto \Gal(L/M) \leq G$ and
 $\beta:H \mapsto \Fix(H) \subseteq L$,
then $\alpha$ and $\beta$ are inverses of each other.
\end{theorem}
\begin{proof}

($\beta \circ \alpha = \Id$). Let $M$ be an intermediate field, and by Theorem \ref{thm:galois-intermediate} we can see that $L/M$ is a Galois Extension. Then if we consider the Galois group $\Gal(L/M)$, we can see that via Theorem \ref{thm:fixed}, $M = \Fix(\Gal(L/M))$. Also, $\alpha(M) = \Gal(L/M)$. Then we look at $(\beta \circ \alpha)(M)$, which is equal to $\beta(\Gal(L/M)) = \Fix(\Gal(L/M))$, but again as see in Theorem \ref{thm:fixed}, we have $\Fix(\Gal(L/M)) = M$. Thus $(\beta \circ \alpha)(M) = M$ and hence we see that $\beta \circ \alpha = \Id$.

($\alpha \circ \beta = \Id$). Now, let $H$ be a subgroup of $G$ and then set $M :=\Fix(H)$. Then $H$ is a subgroup of the Galois group $\Gal(L/M)$. Since $K \subseteq M \subseteq L$, by Theorem \ref{thm:galois-intermediate} we have that $L/M$ is a Galois extension. Thus, by Theorem \ref{thm:equal-subgroup-fixed-points}, since we have $H\leq \Gal(L/M)$ and we have $\Fix(H)=M$, then we can say that $H=\Gal(L/M)=\Gal(L/\Fix(H))$. Thus if we consider $(\alpha \circ \beta)(H) = \alpha(\Fix(H)) = \Gal(L/\Fix(H))= H $, we have $(\alpha \circ \beta)(H) = H$ and so $\alpha \circ \beta = \Id$.

Hence we can see that the functions $\alpha$ and $\beta$ are inverse to each other.
\end{proof}

The next theorem is another important property of Galois extensions. Its proof combines ideas from \cite[Chapter~12]{Stewart} and \cite[p.~58-62]{rotman_galois_1998}. 
\begin{theorem} \label{thm:correspondence-quotient}
    Let $L / K$ be a Galois extension and let $M$ be an intermediate field such that $M /K$ is a Galois extension, then 
    $\Gal(M / K) \cong \Gal(L / K) / \Gal(L / M). $
\end{theorem}

\begin{proof}
	% Theoerm 58, Rotman
	Let $G = \Gal(L/K)$ and $G' = \Gal(M/K)$. Define a map $\psi: G \to G'$ as $\psi(\sigma) = \sigma | _M$ for $\sigma \in G$. We first show that indeed $\psi(\sigma) \in G'$ for all $\sigma \in G$. We only need to prove that $\sigma(M) = M$. $M/K$ is Galois, so by Theorem \ref{thm:fixed}, $M$ is a splitting field for a separable polynomial $f$ over $K$. Let $\alpha_1, \dots, \alpha_n$ be the distinct roots of $f$, then $M = K(\alpha_1, \ldots, \alpha_n)$. But $\sigma$ permutes $\alpha_i$ by Theorem \ref{thm:galois-group-acts-on-zeros} and $\sigma(K) = K$. Therefore $\sigma(M) = \sigma(K(\alpha_i, \ldots, \alpha_n)) = K(\sigma(\alpha_1), \ldots, \sigma(\alpha_n)) = M. $
	
	
	 $\psi$ is clearly a group homomorphism. It is onto due to the normality of $L/K$ and Theorem \ref{thm:monomorphism-extend-automorphism}: each $K$-automorphism $\tau$ of $M$ is also a $K$-monomorphism from $M$ to $L$ and thus there exists a $K$-automorphism $\sigma$ of $L$ such that $\sigma | _M = \tau$. Also, $\sigma \in \operatorname{Ker}(\psi)$ if and only if $\psi(\sigma) = \sigma |_M = \Id$, which is equivalent to $\sigma \in \Gal(L/M)$. Thus $\operatorname{Ker}(\psi) = \Gal(L/M)$. Thus by Theorem \ref{thm:first-iso}, 
	$G / \operatorname{Ker}(\psi) \cong \operatorname{Im}(\psi) = G',$
	and the required result thus follows. 
\end{proof}


% \subsection{Examples of Galois Extensions}
%\begin{example}
%	Consider the separable polynomial $f(t) = t^4-2 $ over $\Q$.
%	
%    $\Q(\sqrt[4]{2}, i)/\Q$ is a Galois Extension. Indeed, consider the separable polynomial $f(t) = t^4-2 $ over $\Q$. The roots of $f$ are $\pm \sqrt[4]{2}$ and $\pm \sqrt[4]{2} i$. We see that the splitting field of $f$ over $\Q$ is $\Q( \sqrt[4]{2},  \sqrt[4]{2}i) = \Q(\sqrt[4]{2}, i)$. We now calculate the Galois group $G = \Gal(\Q(\sqrt[4]{2}, i)/\Q)$. Take $\sigma \in G$, then $\sigma$ is determined by how it maps $\sqrt[4]{2}$ and $i$. 
%    
%    We now consider the subgroup
%    
%
%\end{example}


\begin{example} \label{exm:correspondence}
	
	Let $f(t) = t^3 - 2$ be a separable polynomial over $\Q$ with the set of roots $X = \{ \sq, \zeta \sq, \zeta^2 \sq \}$, where $\zeta = e^{2\pi i / 3}$. 
	
	The splitting field is $F = \Q(\sq, \zeta)$. The minimal polynomial of $\sq$ over $\Q$ is $f$ with degree $3$, and the minimal polynomial of $\zeta$ over $\Q(\sq)$ is $t^2 + t + 1$ with degree $2$. Thus $[F : \Q] = [F : \Q(\sq)] [\Q(\sq) : \Q] = 2 \times 3 = 6$. 
	
	$F / \Q$ is a Galois extension, thus $|G| = [F : \Q] = 6$, where $G = \Gal(F / \Q)$. We know that $G \le S_3$, but $|S_3| = 6$, and thus $G \cong S_3 \cong D_3 = \langle t, r : t^3 = 1, r^2 = 1, tr = rt^{-1} \rangle$. Define $\tau, \rho \in G$ such that 
	$
		\tau(\sq) = \zeta \sq, \, \tau(\zeta) = \zeta; \, \rho(\sq) = \sq, \,  \rho(\zeta) = \zeta ^ 2,
	$
	then it can be verified that $\tau, \rho$ correspond to $t, r \in D_3$ respectively. How $\tau$ and $\rho$ acts on the elements of $X$ can be seen in Figure \ref{fig:example-action}.
	
	We now illustrate the correspondence between the intermediate fields of $F / \Q$ and the subgroups of $G$. For example, we have $\Gal(F / \Q(\zeta)) = \langle \tau \rangle \cong \Z_3 \text{ and } \Fix(\langle \tau \rangle) = \Q(\zeta). $ The full picture can be seen in Figures \ref{fig:example-intermidate-fields} and \ref{fig:example-subgroups}, where corresponding intermediate fields of $F / \Q$ and subgroups of $G$ are drawn in the same color. This example and its figures are adapted from \cite{visual-algebra}. 
	
\end{example}

We now finish this section with a proof of the Fundamental Theorem of Algebra with the help of Galois Theory. Before we begin the proof we recall two important lemmas.

\begin{lemma}\label{lemma:ood-degree-real-root}
    A polynomial of odd degree has a root in the reals.
\end{lemma}

\begin{lemma}\label{lemma:quadratic-extension-in-C}
    Quadratic polynomials with coefficients in $\C$ have roots in $\C$.
\end{lemma}

The above lemma is equivalent to saying that there are no quadratic extensions of $\C$.

Now we have all the tools to prove the Fundamental Theorem of Algebra. The proof that follows has been adapted from Dummit and Foote \cite{foote-algebra}.

\begin{theorem}[Fundamental Theorem of Algebra]
    If $f(t)$ is a non-constant polynomial over $\mathbb{C}$, then there exists $z_0 \in \mathbb{C}$ such that $f\left(z_0\right)=0$.
\end{theorem}

\begin{proof}
    We show that $\C$ is algebraically closed, which is sufficient to show the Fundamental Theorem of Algebra. Let $f \in \R[x]$ with $\partial f = n$, and take $F$ to be a splitting field over $\R$. Then we have that $F(i)$ is a Galois extension over the real numbers. So, define $G:=\Gal(\R/F(i))$ and let $H$ be a $2$-Sylow subgroup of $G$ (see Definition \ref{def:p-group}). Then we have the fixed field of $H$ is an extension of $\R$ which is of an odd degree, and hence by Lemma \ref{lemma:ood-degree-real-root} this is trivial. Thus we can then show that $\Gal(F(i)/\C)$ is a $p$-group with $p=2$, which has a subgroup of any possible order. However, if $\Gal(F(i)/\C)$ is not trivial, this means that there is a quadratic extension of $\C$, which contradicts Lemma \ref{lemma:quadratic-extension-in-C}. Thus we have that $\C$ must be algebraically closed.
\end{proof}
