\section{Ruler-and-Compass Constructions} \label{sec:applications}

The following section is adapted from \cite{Stewart, shape-constructions}. When constructing shapes, the Greeks, specifically Plato, looked at `perfect' geometric figures - those which can be constructed with only straight lines and a compass. This is to see if certain shapes and figures can be drawn without approximations. Euclid set out five main axioms to drawing these figures:

\begin{enumerate}
    \item One can draw a straight line segment between any two points.
    \item One can extend any line segment infinitely.
    \item Given a point and a distance $r$, one can construct a circle centered at that point, with radius $r$.
    \item Any two right angles are congruent.
    \item Given a line $L$ and a point $P$, one can construct a line through $P$ that is parallel to $L$. 
\end{enumerate}

\subsection{Constructable Numbers}

Now we can begin a formal definition, so now we construct the plane $\R^2$ which is isomorphic to $\C$. 
\begin{definition}
Let points be denoted by $z_1,z_2 \in \C$ and we let $0\leq r \in \R$.
    \begin{itemize}
        \item $L(z_1,z_2)$ is the (infinite) straight line joining $z_1$ to $z_2$ given that $z_1 \neq z_2$. 
        \item $C(z_1,r)$ is the circle around $z_1$ with radius $r$.
    \end{itemize}
    For the following we take $n \in \N$, and define $\mathcal{P}_n,\mathcal{L}_n,\mathcal{C}_n$ as the sets of $n$-constructable points, lines and circles respectively.
    \begin{itemize}
        \item $\mathcal{P}_0=\{0,1\}$; $\mathcal{P}_{n+1}=\{z\in \C $: $z$ lies on two distinct lines in $\mathcal{L}_{n+1}$ or it lies on one line in $\mathcal{L}_{n+1}$ and one circle in $\mathcal{C}_{n+1}$ or $z$ lies on two distinct circles in $\mathcal{C}_{n+1}$\};
        \item $\mathcal{L}_0=\varnothing$; $\mathcal{L}_{n+1}=\{L(z_1,z_2) : z_1,z_2\in \mathcal{P}_n\}$;
        \item $\mathcal{C}_0=\varnothing$; $\mathcal{C}_{n+1}=\{C(z_1,|z_2-z_3|:z_1,z_2,z_3\in \mathcal{P}_n\}$.
    \end{itemize}
\end{definition}

\begin{definition}\label{def:constructable-point}
    A point $z \in \C$ is called \textit{constructable} if there is a finite sequence of points:
    $z_0=0,z_1=1,z_2,z_3,...,z_k=z$, where $z_{j+1}$ lies in one of the following:
    \begin{itemize}
        \item On the intersection of two straight lines constructed between two sets of two points, i.e.
        $z_{j+1}\in L(z_{j_1},z_{j_2})\cap L(z_{j_3},z_{j_4})$;
        \item On the intersection of a circle and a line, i.e.
        $z_{j+1}\in L(z_{j_1},z_{j_2})\cap C(z_{j_3},|z_{j_4}-z_{j_5})|$;
        \item On the intersection of two circles, i.e.
        $z_{j+1}\in C(z_{j_1},|z_{j_2}-z_{j_3})|\cap C(z_{j_4},|z_{j_5}-z_{j_6})|$;
    \end{itemize}
    where $j_i\in \N$.
\end{definition}

\begin{lemma}\label{thm:point-subset}
    For any $n \in \N$, 
    $\mathcal{P}_n \subseteq \mathcal{P}_{n+1}$. 
\end{lemma}

%\begin{proof}
%    The proof is trivial, clearly all the points which lie in  $\mathcal{P}_n$ have to lie in $\mathcal{P}_{n+1}$.
%\end{proof}

\begin{theorem}
    A point is called constructable if and only if $z\in \mathcal{P}_n$ for some $n\in \N$. 
\end{theorem}

\begin{proof}  
    `$\Longrightarrow$'. Let $z \in \C$ be a constructable point, then by Definition \ref{def:constructable-point}, it is clear that $z=z_k\in \mathcal{P}_k$.
    
    `$\Longleftarrow$'. On the other hand, if $z \in \mathcal{P}_k$, then we can find a sequence $z_j \in \mathcal{P}_j$ for $0\leq j \leq k$ which satisfies Definition \ref{def:constructable-point}.
\end{proof}

We now define the `Pythagorean Closure' which will help us create a link between construction and Galois Theory.

\begin{definition}[Pythagorean Closure]
    $\Q^{py}$ is called the `Pythagorean Closure' of $\Q$, which is the union of all subfields, denoted $K$ where
    $\Q = K_0 \subseteq K_1 \subseteq ... \subseteq K_n=K$ such that $[K_{j+1}:K_j]\leq 2$.
\end{definition}

Here we can see a tower of field extensions, which gives us a hint that we may be able to use fields and extensions to solve construction problems.

The next lemma will help us to prove a theorem which will show the link between the two.

\begin{lemma}\label{lemma:conjugate-in-Pn}
    Let $z \in \mathcal{P}_n$, then its complex conjugate $\overline{z} \in \mathcal{P}_{n+1}$.
\end{lemma}

\begin{proof}  
    A visual proof is also included in the appendix (see Figure \ref{fig:case-1-conjugate} and Figure \ref{fig:case-2-conjugate}).
    
    Let us show that we can construct the complex conjugate using the points $0, 1, z\in \mathcal{P}_{n+1}$.
    We start with the trivial case where either $z=1$ or $\overline{z}=1$. Here the point has already been constructed, and since $\overline{z}=1\in \mathcal{P}_0$ this implies $\overline{z}=1\in \mathcal{P}_n$ by Lemma \ref{thm:point-subset}.


    Now, let us say $z\in \mathcal{P}_n$ with $z\neq1$. Then we consider the circle $C(z,|z-1|)$, and the line $L(0,1)$. We have two cases here, either the circle meets the line at exactly one point, $1$, or the circle meets the line in exactly two points, $1$ and $s$.

    \textbf{Case One:} If $C(z,|z-1|)$ and $L(0,1)$ meet at exactly one point, then the line $L(1,z)$ must be perpendicular to $L(0,1)$. Then by drawing the circle $C(1,|z-1|)$, we see that it meets the line $L(1,z)$ at two points, one is $z$ and the other is $\overline{z}$.

    \textbf{Case Two:} If $C(z,|z-1|)$ and $L(0,1)$ meet at two points, then one of them must be $1$ and we shall call the other $z_1\neq1$. Then construct two more circles $C_1=C(1,|z-1|)$ and $C_2=C(z_1,|z-1|)$. Then $C_1$ and $C_2$ have two points of intersection, one being $z$ and the other $\overline{z}$.

    Thus we have all points lie in $\mathcal{P}_n$ and thus $\overline{z} \in \mathcal{P}_{n+1}$.
\end{proof}

\begin{theorem}\label{thm:power-of-two-construction}
    If $z$ is constructable, then $[\Q(z):\Q]$ is a power of two.
\end{theorem}

\begin{proof}  
   Consider we already have a set of constructable points $X_k = \{0,1,z_1,...,z_k\}$ (where $X_0=\{0,1\}$), and we have another point, $z=z_{k+1}$ which can be constructed through the intersections of lines and circles with regards to these points. Then, algebraically, we have a linear or quadratic equation, where $z_{k+1}$ is a solution. This is due to the fact that a line has the equation $y=mx+c$ and a circle has equation $(x-a)^2+(y-b)^2=r^2$, where $z=x+iy$ and $r=x^2+y^2$. Thus, the new point extends the field by order one (if linear) or at most two (if quadratic), so $[\Q(z_{k+1}):\Q(X_k)] \in \{1,2\} \forall k \in \N$. Thus, by Theorem \ref{thm:tower-theorem}, we see that $[\Q(z):\Q] = \prod_{i=0}^k [\Q(z_{i+1}):\Q(X_i)]$ and therefore, since each $[\Q(z_{k+1}):\Q(X_k)] \in \{1,2\}$, we have that $[\Q(z):\Q] = 2^k$ for some $k \in \N$.
\end{proof}


Now we have the key results to help us prove the following theorem, where the proof here is adapted from \cite{constructions-and-galois}.

\begin{theorem}
    A point $z \in \C$ is constructable if and only if $z \in \Q^{py}$.
\end{theorem}

\begin{proof}
    `$\Longrightarrow$'. Let $P(n)$ be the statement $\mathcal{P}_n \subseteq \Q^{py}$.
     We have that $\mathcal{P}_0 = \{0,1\}$, which clearly $\{0,1\}\subseteq\Q^{py}$, so $P(0)$ holds true.
    Now assume that $P(k)$ holds true, and now consider $P(k+1)$. So we let $z \in \mathcal{P}_{k+1}$. Then we have three cases:
    \begin{itemize}
        \item \textbf{Case One:} Intersection of lines. If we have $z$ is the intersection of lines from points $z_1, z_2, z_3, z_4 \in \mathcal{P}_{k+1}$, then we can form 2 simultaneous equations:
        $z = az_1+(1-a)z_2$ and $z = bz_3+(1-b)z_4$. By algebraic manipulation (shown in the appendix) we can find formulae for $a$ and $b$ in terms of the points $z_1,z_2,z_3$ and $z_4$, so we know that $a,b \in \Q^{py}$, and since $z_1,z_2,z_3,z_4\in \Q^{py}$ by our assumption, we have that $z\in \Q^{py}$ also, and hence in this case $\mathcal{P}_n \subseteq \Q^{py}$.
        \item \textbf{Case Two:} Intersection of a line and circle. Here we look at $z \in L(z_1,z_2)\cap C(z_3,|z_4-z_5|$ where $z_i \in \mathcal{P}_k$ for $1\leq i \leq 5$. Then, again we get two simultaneous equations:
        $z=az_1+(1-az_2)$ and $z=z_3+|z_4-z_5|e^{i\theta}$ where $a$ and $\theta$ are to be found and $a,\theta \in \R$. Again, through manipulation, we can find that $a$ can be found through a quadratic equation with coefficients in $\Q^{py}$ and so $a \in \Q^{py}$ which implies that $z\in \Q^{py}$.
        \item \textbf{Case Three:} Intersection of circles. So we say that $z \in C(z_1,|z_2-z_3|) \cap C(z_4,|z_5-z_6|)$ where $z_i \in \mathcal{P}_n$ for $i = 1,2,...,6$. In this case we again have 2 simultaneous equations: $z=z_1+|z_2-z_3|e^{i\theta}$ and $z=z_4+|z_5-z_6|e^{i\phi}$, for $\theta,\phi \in \R$. Here we can take conjugates and show that $(z-z_1)(\overline{z}-\overline{z_1})=|z_2-z_3|^2$. From this we can solve for $z$ and $\overline{z}$ and thus see that $z \in \Q^{py}$.
    \end{itemize}
Therefore, $P(k+1)$ holds, and thus $\mathcal{P}_n \subseteq \Q^{py}$. 

`$\Longleftarrow$'. Now we show that for any $z \in Q^{py}, \exists n \in \N$ such that $z \in \mathcal{P}_n$.
So let $z \in \Q^{py}$, and by the definition of $\Q^{py}$ we know that  $z\in \Q^{py}$ if and only if $z \in K_n$, the highest field in our tower. Thus, there must be a sequence of points, $0,1,z_0,z_1,...,z_k=z$, where $z_i \in K_i$ for $0\leq i \leq n$, such that $z_{i+1} \in \Q(0,1,...,z_i,\alpha)$ for $\alpha^2 \in \Q(0,1,...,z_i)$. This follows from the definition of $\Q^{py}$ as it is the union of subfields which contain the $z_i$. Then we have that $z_i$ is constructable, thus $z_{i+1}$ must also be constructable. So all points in this sequence are constructable, thus $z$ is constructable and therefore for $z \in \Q^{py}$ we have that $z \in \mathcal{P}_n$ for some $n \in \N$.

Thus we have shown that $z$ is constructable if and only if $z\in \Q^{py}.$
\end{proof}



\subsection{Impossibility Proofs}

There are many shapes and objects that are not possible to construct using just a ruler and compass. One of the most famous examples of this, is the construction of a square with the same area of a circle. In this section, we look at such impossible constructions, and show why this is impossible.

\begin{theorem}
    The regular heptagon cannot be constructed with ruler and compass.
\end{theorem}

\begin{proof}
    If the regular heptagon were to be constructable, then this means the point $z=e^{2i\pi/7} \in \Q^{py}$. This is one of our complex roots of unity, and this means that $z^7+z^6+z^5+...+z=0$. Since $z$ is a root of unity, $z^7=1 \implies z^7 - 1 = 0$. We can factorise this as $z^7-1=(z-1)(z^6+z^5+...+z+1)$. Then by Theorem \ref{thm:irreducible-prime-polynomial}, we know that since 7 is prime, $z^6+z^5+...+z+1$ is irreducible over $\Q$ and thus $[\Q(z):\Q]$ has degree 6. We know that six is not a power of two, which means this contradicts Theorem \ref{thm:power-of-two-construction}, therefore a heptagon is not constructable.
\end{proof}

We now prove the impossibility of doubling the volume of a cube using only a compass and a straight edge.

\begin{theorem}
    The cube cannot be doubled by ruler-and-compass construction.
\end{theorem}

\begin{proof}  
    We can first see that to double the volume of a cube, we would need to increase the edge length by a scale factor of $\sqrt[3]{2}$. Thus we can formulate this problem to see if constructing $\sqrt[3]{2}$ is possible, that is to say $z = \sqrt[3]{2} \in \Q^{py}$.

    So let us assume towards a contradiction that $\sqrt[3]{2} \in \Q^{py}$ with $\mu$ being its minimal polynomial over $\Q$. Then, by Theorem \ref{thm:power-of-two-construction}, we have that the degree of $\mu = 2^k$ for some $k\in \N \cup \{0\}$.
    Then clearly $z^3=2$, so $z^3-2=0$ we must have that $\mu$ divides $z^3-2$, which is irreducible over $\Q$, thus the field extension $\Q(\sqrt[3]{2}):\Q$ has order $3$, which is not a power of two, which contradicts Theorem \ref{thm:power-of-two-construction}, and thus $z = \sqrt[3]{2} \notin \Q^{py}$.
\end{proof}


`Squaring the circle' is a problem posed by the Greeks which asked if it was possible to construct a square with the same area as a circle, using only a compass and a straight edge. The following theorem shows the solution to the posed problem.

\begin{theorem}
    The circle cannot be squared using a ruler-and-compass construction.
\end{theorem}

\begin{proof}  
    
    It is sufficient to show that $\sqrt{\pi}\in \C$ is constructable from our initial set of points $\mathcal{P}_0$. From this we could construct $\pi$. Thus, we would have that $[\Q(\pi):\Q]$ is a power of 2, which would imply that $\pi$ is algebraic over $\Q$. However, as seen by Lindemann's proof (which we include in Appendix \ref{sec:construction}), $\pi$ is not algebraic over $\Q$ and thus it is impossible to construct a square with the same area as a circle, using nothing but a compass and a straight edge.
\end{proof}
